---
title: "headsup_onfarm"
author: "Ryan Hamilton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
setwd("C:/Heads Up") #sets working directory

knitr::opts_chunk$set(echo = TRUE)
```

### Start Clean
```{r eval = TRUE, echo = FALSE, results = 'hide'}
rm(list = ls(all=TRUE)) # removes all variables in the global environment so you start fresh

options(scipen = 999) # stops anything from being in scientific notation
```

### Load dependencies, and define some basic functions
```{r load dependencies, echo = FALSE,  message = FALSE}

#tidyverse/miscellaneous packages
library("tidyverse")
library("plyr")
library("dplyr")
library("stringr")
library("latticeExtra")
library("knitr")
library("gridExtra")
library("scales")

#visualization/ggplot packages
library("ggplot2")
library("ggpubr")
library("ggridges")
library("ggsignif")
library("RColorBrewer")
library("ggrepel")

#mapping packages
library("purrr")
library("terra")

#statistical analysis packages
library("rcompanion")
library("FSA")
library("rstatix")
library("car")
library("matrixStats")
library("emmeans")
library("multcomp")
library("lme4")

#trims leading and trailing spaces
trim <- function (x) gsub("^\\s+|\\s+$", "", x)

#converts factors to numeric values
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}

#creates standard error function from sd and sqrt of number of observations
se <- function(x) sd(x,na.rm=T)/sqrt(length(x))

```

### Read in dataset and parses variables
```{r include=FALSE}
headsup_strip<-read.csv(file="C:/Heads Up/headsup_strip.csv",header=T) #reads in dataset
facs<-c("year","location","rep","trt","product") #assigns factor names to a vector
headsup_strip[facs]<-lapply(headsup_strip[facs],factor) #uses the char vector to parse variables by col name
headsup_strip<-headsup_strip %>% mutate(yield_kgha = (yield_bua*67.25))
str(headsup_strip) #dataframe summary by string
```

### Let's look at the data and do some resid diagnostics
```{r echo=FALSE}
(lmod<-lm(yield_kgha~product+location,data=headsup_strip)) #sets up a simple linear model 
plot(lmod) #no linear relationship evident bw product and yield
joint_tests(lmod)
Anova(lmod)
drop1(lmod)
#should we consider transforming the response variable (yield)?
#(boxcox(lmod)) #no transformation needed, thankfully

```

### Mixed models and summary stats by independent variable = product
```{r include=FALSE}

#yield vs. product model -- no sig. diff
lmod_mix<-lmer(yield_kgha~product+(1|year:location:rep),data=headsup_strip) #model optimized for split plot/repeated measures
#lmod_mix<-lmer(yield_kgha~product+(1|location)+(1|year)+(1|rep),data=headsup_strip) #mixed effects model for RCBD
summary(lmod_mix) #model summary
drop1(lmod_mix) #parameter selection
joint_tests(lmod_mix) #emmeans gtype 3 ANOVA
Anova(lmod_mix,type=3) #car type 3 ANOVA 
lsmeans<-emmeans(lmod_mix,"product") #calculates lsmeans
cld(lsmeans,Letters=letters,alpha=.05) #default is Tukey
pairs(lsmeans) #all pairwise comparisons

shapiro_test(headsup_strip$yield_kgha) #not normal
levene_test(headsup_strip,yield_kgha~product) #very homogeneous
length(headsup_strip$yield_kgha) #n=98, so we can do type I ANOVA on product vs. yield relationships

#overall yield summary by product
yield_prod<-ddply(headsup_strip, .(product), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd(as.numeric(yield_kgha),na.rm=T)/sqrt(length(yield_kgha)))

#overall yield summary by product
yield_prod<-ddply(headsup_strip, .(location), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd(as.numeric(yield_kgha),na.rm=T)/sqrt(length(yield_kgha)))

#yield summary by year, location, and product
yield_prod2<-ddply(headsup_strip, .(year,location,product), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))

#KW and Dunn's tests for pairwise yield comparisons within site
(headsup_strip %>% group_by(location) %>% dunn_test(yield_kgha~product)) #ok, one sig. diff at one site in one year

```

### Mixed models and summary stats by independent variable = location
```{r include=FALSE}
#yield vs. location model -- def sig. diff between locations
lmod_mix2<-lmer(yield_kgha~location+(1|product)+(1|year:location:rep),data=headsup_strip) #model optimized for split plot/repeated measures
#lmod_mix2<-lmer(yield_kgha~product+location+(1|year)+(1|rep),data=headsup_strip) #mixed effects model
summary(lmod_mix2) #model summary
drop1(lmod_mix2) #parameter selection
joint_tests(lmod_mix2) #emmeans gtype 3 ANOVA
Anova(lmod_mix2,type=3) #car type 3 ANOVA 
lsmeans2<-emmeans(lmod_mix2,"location") #calculates lsmeans
cld(lsmeans2,Letters=letters,alpha=.05) #default is Tukey
pairs(lsmeans2)

shapiro_test(headsup_strip$yield_kgha) #recall that this is NOT normal
levene_test(headsup_strip,yield_kgha~location) #not at all homogeneous
#length(headsup_strip$yield_kgha) #n=98, but we can't perform type I ANOVA bc 2 assumptions are violated - KW ONLY!

#KW and Dunn's tests for pairwise yield comparisons between sites
#these tests provide exact p-values for the cld-type output above
(headsup_strip %>% group_by(year) %>% dunn_test(yield_kgha~location)) #ok, one sig. diff at one site in one year

yield_year<-ddply(headsup_strip, .(year), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))

#overall site-level yield summary by product for comparison of mean yield by location - KW only
yield_loc<-ddply(headsup_strip, .(location), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))

#yield summary by year, location, and product - KW only
yield_loc2<-ddply(headsup_strip, .(year,location,product), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))
```

### Mixed models and summary stats by independent variable = year
```{r include=FALSE}
#yield vs. year -- soybean yield is not sig. diff between years
lmod_mix3<-lmer(yield_kgha~year+(1|product)+(1|location:rep),data=headsup_strip) 
#lmod_mix3<-lmer(yield_kgha~product+(1|location)+year+(1|rep),data=headsup_strip) #mixed effects model
summary(lmod_mix3) #model summary
drop1(lmod_mix3) #parameter selection
joint_tests(lmod_mix3) #emmeans gtype 3 ANOVA
Anova(lmod_mix3,type=3) #car type 3 ANOVA 
lsmeans3<-emmeans(lmod_mix3,"year") #calculates lsmeans
cld(lsmeans3)
pairs(lsmeans3) #default is Tukey

#shapiro_test(headsup_strip$yield_kgha) #recall that this is not normal
(levene_test(headsup_strip,yield_kgha~year)) #not at all homogeneous
#length(headsup_strip$yield_kgha) #n=98, but we can't perform type I ANOVA bc 2 assumptions are violated

#KW and Dunn's tests for pairwise yield comparisons within sites
#these tests provide exact p-values for the cld-type output above
(headsup_strip %>% #group_by(year) %>% 
    dunn_test(yield_kgha~product)) #no sig. diff in either year -- exact p-values

#overall experiment-level yield summary by year - KW only
yield_year<-ddply(headsup_strip, .(year), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))

#yield summary by year, location, and product - KW only
yield_year2<-ddply(headsup_strip, .(year,product), summarize,mean_yield=mean(as.numeric(yield_kgha),na.rm=T),se_yield=sd((as.numeric(yield_kgha))/sqrt(length(yield_kgha)),na.rm=T))
```

### summary stats for plotting -- turn on if needed
```{r eval=FALSE, include=FALSE}
mean(strip_base$yield_kgha,na.rm=T)
sd(strip_base$yield_kgha,na.rm=T)
se(strip_base$yield_kgha)

mean(strip_plot$yield_kgha,na.rm=T)
sd(strip_plot$yield_kgha,na.rm=T)
se(strip_plot$yield_kgha)

mean(strip_plot$yield_kgha,na.rm=T)-mean(strip_base$yield_kgha,na.rm=T)
```

### Piano plot
```{r eval=FALSE, message=TRUE, warning=TRUE, include=FALSE}
strip_means<-read.csv("C:/Heads Up/strip_plot.csv",header=T) #reads in data
facs<-c("X","year","location","rep","trt","product") #string of column headers
strip_means[facs]<-lapply(strip_means[facs],factor) #assigns factors
str(strip_means) #views summary of the strings

strip_base<-filter(strip_means,product=="BASE") #subsets the base yield response means
strip_plot<-filter(strip_means,product=="BASE+HEADSUP") #subsets base+headsup for piano-plotting
strip_plot<-filter(strip_plot,mean_diff!="NA")#removes lines with NAs for plotting
attach(strip_plot) #brings the df into local environment
leg<-expression(paste("kg/ha"))

  piano2<-ggplot(strip_plot, aes(x=X,y=mean_diff)) + 
    geom_bar(aes(fill=mean_diff),position="dodge",stat="identity",width=0.8,fill="orange",color="darkorange") +
    scale_y_continuous(limits=c(-1000,1000,by=200)) +
    aes(x=fct_reorder(X,mean_diff)) +
    labs(title=element_text("On-farm strip trials (2022-2023)"),y=expression(paste("Relative yield difference (kg/ha)")), x=element_text("On farm trials (2022-2023): Differential soybean yield response by experimental replicate (n = 44)")) +
    theme_classic() +
    theme(axis.text.y=element_text(color="black",size=14)) +
    theme(axis.text.x=element_blank()) + 
    theme(axis.ticks.x=element_blank()) +
    theme(axis.title.y=element_text(color="black",size=18)) +
    theme(axis.title.x=element_text(color="black",size=18)) +
    theme(plot.title=element_blank()) +
    geom_hline(yintercept=mean(strip_plot$mean_diff,na.rm=T)) + #numeric difference between treatment means (basically 0)
    geom_hline(linetype="dashed",yintercept=mean(strip_plot$mean_diff,na.rm=T)+(1.96*se(strip_plot$yield_kgha))) + #line at the upper 95% CI limit of the mean yield
    geom_hline(linetype="dashed",yintercept=mean(strip_plot$mean_diff,na.rm=T)-(1.96*se(strip_plot$yield_kgha))) #line at the lower 95% CI limit of the mean yield
  piano2
  piano_leg2<-piano2
  piano_noleg2<-piano2 +
    theme(legend.position="none")
 detach(strip_plot)
#ggsave("strip_leg.png",path="C:/Heads Up",plot=piano_leg,width=12,height=8,units="in",dpi=320)
ggsave("strip_noleg2.png",path="C:/Heads Up",plot=piano_noleg2,width=12,height=8,units="in",dpi=600)
```

### Yield plots
```{r warning=FALSE}
#let's plot this real quick
strip22<-headsup_strip %>% filter(year=="2022") #subsets for plotting by year
strip23<-headsup_strip %>% filter(year=="2023")

    farm_plot22<-ggplot(data=na.omit(strip22[,c("yield_kgha","location","product")]),aes(x=location,y=(yield_kgha),fill=product)) +
      geom_boxplot() +
      ylim(0,6000) +
      scale_fill_hue(labels = c("Base", "Base +\nHeads Up")) +    
      labs(title="2022",y=expression(paste("Soybean yield (kg/ha)"))) +
      theme_classic() +
      theme(axis.text.x=element_text(angle=45,vjust=0.75,hjust=0.75,color="black")) +
      theme(legend.text=element_text(size=12)) +
      theme(legend.title=element_blank()) +    
      theme(axis.title.x = element_blank()) +
      theme(axis.text.x = element_text(size=16)) +
      theme(axis.text.y = element_text(size=12)) +
      theme(axis.line.x = element_line(color="black", size = 0.8),axis.line.y = element_line(color="black",size=0.8)) +
      theme(axis.title.y = element_text(color="black",size=18,face="bold")) +
      theme(plot.title=element_text(color="black",size=18,face="bold",hjust=0.5)) +
      stat_kruskal_test(aes(label=sprintf("P = %5.2f",as.numeric(..p.adj..)),p.adjust.method="none"),size=4,label.y=1,vjust=-1,angle=45)
      
    farm_plot22 
    
ggsave("farm_plot22.png",path="C:/Heads Up",plot=farm_plot22,width=8,height=8,units="in",dpi=600)

    farm_plot23<-ggplot(data=na.omit(strip23[,c("yield_kgha","location","product")]),aes(x=location,y=(yield_kgha),fill=product)) +
      geom_boxplot() +
      ylim(0,6000) +
      scale_fill_hue(labels = c("Base", "Base +\nHeads Up")) +    
      labs(title="2023",y=expression(paste("Soybean yield (kg/ha)"))) +
      theme_classic() +
      theme(axis.text.x=element_text(angle=45,vjust=0.75,hjust=0.75,color="black")) +
      theme(legend.text=element_text(size=12)) +
      theme(legend.title=element_blank()) +    
      theme(axis.title.x = element_blank()) +
      theme(axis.text.x = element_text(size=16)) +
      theme(axis.text.y = element_text(size=12)) +
      theme(axis.line.x = element_line(color="black", size = 0.8),axis.line.y = element_line(color="black",size=0.8)) +
      theme(axis.title.y = element_blank()) +
      theme(plot.title=element_text(color="black",size=18,face="bold",hjust=0.5)) +
      stat_kruskal_test(aes(label=sprintf("P = %5.2f",as.numeric(..p.adj..)),p.adjust.method="none"),size=4,label.y=1,vjust=-1,angle=45)
      
    farm_plot23 
    
ggsave("farm_plot23.png",path="C:/Heads Up",plot=farm_plot23,width=8,height=8,units="in",dpi=600)
```